---
title: "Analyzing land productivity in Madison (WI)"
author: "Harald Kliems"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data
The Madison Open Data portal has property tax information by parcel: https://data-cityofmadison.opendata.arcgis.com/datasets/tax-parcels-assessor-property-information/data

Data was manually downloaded as a shapefile and then saved as an `rds` object for faster loading.

```{r}
library(tidyverse)
library(sf)
library(ggplot2)
```


```{r eval=FALSE, include=FALSE}
parcels <- read_sf("data/Tax_Parcels__Assessor_Property_Information_.shp")
saveRDS(parcels, "data/parcels.rds")
```
The dataset is large, with over 80,000 observations and 144 variables. It does not contain information on ownership.

```{r}
parcels <- readRDS("data/parcels.rds")
str(parcels)
```
I don't see a data dictionary, but most of the variable names are self-explanatory. Some interesting stuff:
- There are variables for noise from streets, railroads, and the airport. But it's unclear what the values mean. I initially thought they were decibels from noise measurements, but the values are other 0, or 60 (airport), 61 (street, railroad)

```{r}
parcels %>% 
  filter(NoiseRailr > 0) %>% 
  select(Address, NoiseRailr) %>% 
  arrange(desc(NoiseRailr))
```
- There is info on building styles, with 29 different classifications:
```{r}
unique(parcels$HomeStyle)
```
# Property tax per sqft
Add a new variable for `NetTaxes` per sq ft.

```{r}
parcels2 <- parcels %>% 
  mutate(tax_per_sqft = NetTaxes / LotSize)
```



```{r}
parcels2 %>% 
  mutate(Zoning1 = fct_reorder(Zoning1, tax_per_sqft, median)) %>% 
  group_by(Zoning1) %>% 
  ggplot(aes(Zoning1, tax_per_sqft)) +
  geom_boxplot() +
  scale_y_log10() +
  coord_flip()
```
Unfortunately, the `Zoning1` field doesn't always contain the current zoning code. For instance, there are many codes starting with `HIS-`. I firstx thought these may represent hstorical zoning codes, but it looks like they are the city's designated historic districts. https://www.cityofmadison.com/dpced/planning/local-historic-districts/1601/ In most cases the actual zoning district then is in the `Zoning2` field, and in a few cases, you have to go to `Zoning3`.

```{r}
parcels3 <- parcels2 %>% 
  mutate(zoning_clean = case_when(str_detect(Zoning1, "HIS-") & !str_detect(Zoning2, "HIS-") ~ Zoning2,
                                  str_detect(Zoning1, "HIS-") & str_detect(Zoning2, "HIS-") ~ Zoning3,
                                  TRUE ~ Zoning1))
```



```{r}
library(mapdeck)
mapdeck(style = 'mapbox://styles/mapbox/dark-v9', pitch = 45) %>% 
  add_polygon(parcels3, elevation = "tax_per_sqft",
             elevation_scale = 1000000)
```

# Issues
- Condos are difficult to handle for the tax per sqft of lot size because their lot size is provided as 0.